Follow-up Samples

### 2016-12-05 submission/manuscript_outline_20161024.Rmd
abstract update v 3

Colorectal cancer (CRC) continues to be a worldwide health problem with early detection being used as a key component in mitigating deaths due to the disease.  Previous research suggests that the bacterial microbiome can be used as a biomarker for colorectal cancer (CRC).  These reports have mostly focused on investigating how the bacterial microbiome is used at a single point in time to predict disease.  In this study we assessed whether a model built with bacterial microbiome data could accurately predict adenoma or carcinoma and adjust positive probability for either, after surgical treatment.  This model was tested on a ```r sum(initial_cm$ref_yes)``` person group that included samples before and after treatment to allow for the assessment of how the model adjusts risk after treatment. The model chosen had a 10-fold cross validated AUC of ```r format(model_test_AUC_summary["middle", "ROC_cv"], digits = 3)``` and a test set AUC of ```r format(model_test_AUC_summary["middle", "AUC"], digits = 3)```.  There was a significant decrease in probabliity of lesion from the model after treatment for both the adenoma (P-value = ```r format(IF_model_wilcox_paired_summary['all_adenoma', 'BH_correction'], scientific = TRUE, digits = 3)```) and carcinoma (P-value = ```r format(IF_model_wilcox_paired_summary['carcinoma_only', 'BH_correction'], scientific = TRUE, digits = 3)```) group.  The top 5 most important operational taxonomic units (OTUs) for prediction of lesion classified to ```r paste(rf_lowest_ID[1], " (", names(rf_lowest_ID)[1], ")", sep = "")```,  ```r paste(rf_lowest_ID[2], " (", names(rf_lowest_ID)[2], ")", sep = "")```, ```r paste(rf_lowest_ID[3], " (", names(rf_lowest_ID)[3], ")", sep = "")```, ```r paste(rf_lowest_ID[4], " (", names(rf_lowest_ID)[4], ")", sep = "")```, and ```r paste(rf_lowest_ID[5], " (", names(rf_lowest_ID)[5], ")", sep = "")```.  The prediction of lesion on initial samples had a sensitivity of ```r format(confusion_summary["Sensitivity", "initial"]*100, digits = 3)```% (```r paste(initial_cm["pred_yes", "ref_yes"], "/", sum(initial_cm$ref_yes), sep = "")```) while follow up it was ```r format(confusion_summary["Sensitivity", "followup"]*100, digits = 3)```% (```r paste(followup_cm["pred_yes", "ref_yes"], "/", sum(followup_cm$ref_yes))```).  Overall, this model could accurately predict those with an adenoma or carcinoma and lower probability of having a lesion based on follow up samples after treatment.

>>>>

Colorectal cancer (CRC) continues to be a worldwide health problem with early detection being used as a key component in mitigating deaths due to the disease.  Previous research suggests that the bacterial microbiome can be used as a biomarker for colorectal cancer (CRC).  These reports have mostly focused on investigating how the bacterial microbiome is used at a single point in time to predict disease.  In this study, we assessed whether a model built with bacterial microbiome data could accurately predict adenoma or carcinoma and adjust positive probability for lesion (adenoma or carcinoma) after surgical treatment.  This model was tested on a ```r sum(initial_cm$ref_yes)``` person group that included samples before and after treatment to allow for the assessment of how the model adjusts risk after treatment. The model chosen had a 10-fold cross validated AUC of ```r format(model_test_AUC_summary["middle", "ROC_cv"], digits = 3)``` and a test set AUC of ```r format(model_test_AUC_summary["middle", "AUC"], digits = 3)```.  : For the follow up samples our Random Forest model significantly decreased the positive probability of lesion compared to the initial samples for both adenoma (P-value = ```r format(IF_model_wilcox_paired_summary['all_adenoma', 'BH_correction'], scientific = TRUE, digits = 3)```) and carcinoma (P-value = ```r format(IF_model_wilcox_paired_summary['carcinoma_only', 'BH_correction'], scientific = TRUE, digits = 3)```).  The top 5 most important operational taxonomic units (OTUs) for prediction of lesion classified to ```r paste(rf_lowest_ID[1], " (", names(rf_lowest_ID)[1], ")", sep = "")```,  ```r paste(rf_lowest_ID[2], " (", names(rf_lowest_ID)[2], ")", sep = "")```, ```r paste(rf_lowest_ID[3], " (", names(rf_lowest_ID)[3], ")", sep = "")```, ```r paste(rf_lowest_ID[4], " (", names(rf_lowest_ID)[4], ")", sep = "")```, and ```r paste(rf_lowest_ID[5], " (", names(rf_lowest_ID)[5], ")", sep = "")```.  In the test set, the prediction of lesion for the initial samples had a sensitivity of ```r format(confusion_summary["Sensitivity", "initial"]*100, digits = 3)```% (```r paste(initial_cm["pred_yes", "ref_yes"], "/", sum(initial_cm$ref_yes), sep = "")```) while follow up it was ```r format(confusion_summary["Sensitivity", "followup"]*100, digits = 3)```% (```r paste(followup_cm["pred_yes", "ref_yes"], "/", sum(followup_cm$ref_yes))```).  Overall, this model could accurately predict those with an adenoma or carcinoma and lower probability of having a lesion based on follow up samples after treatment.

### 2016-12-05 submission/manuscript_outline_20161024.tex
abstract update v 3

Colorectal cancer (CRC) continues to be a worldwide health problem with
early detection being used as a key component in mitigating deaths due
to the disease. Previous research suggests that the bacterial microbiome
can be used as a biomarker for colorectal cancer (CRC). These reports
have mostly focused on investigating how the bacterial microbiome is
used at a single point in time to predict disease. In this current study
we attempt to combine specific physical characteristics, patient
history, Fecal Immunochemical Test (FIT), and the bacterial microbiome
to build the best model for classifying those with and without lesion
(adenoma or carcinoma). This model was tested on a \texttt{66} person
group that included samples before and after treatment to allow for the
assessment of how the model adjusts risk after treatment. Using NMDS,
the thetayc distance between initial and follow up showed a significant
difference (P-value = \texttt{0.001}) for the carcinoma group only. The
model chosen had a 10-fold cross validated AUC of \texttt{0.819} and a
test set AUC of \texttt{0.815}. There was a significant decrease in
probabliity of lesion from the model after treatment for both the
adenoma (P-value = \texttt{8.48e-07}) and carcinoma (P-value =
\texttt{5.96e-08}) group. The top 5 most important operational taxonomic
units for this model were \texttt{Lachnospiraceae\ (Otu000013)},
\texttt{Escherichia/Shigella\ (Otu000018)},
\texttt{Ruminococcaceae\ (Otu000020)},
\texttt{Ruminococcus\ (Otu000017)}, and
\texttt{Porphyromonas\ (Otu000153)}. For prediction on samples with
follow up the initial sensitivity was \texttt{98.5}\% (\texttt{65/66})
and follow up sensitivity was \texttt{100}\% (\texttt{1\ /\ 1}).
Overall, this model can accurately predict and subsequently respond to
treatment by lowering the probability of lesion.

>>>>

Colorectal cancer (CRC) continues to be a worldwide health problem with
early detection being used as a key component in mitigating deaths due
to the disease. Previous research suggests that the bacterial microbiome
can be used as a biomarker for colorectal cancer (CRC). These reports
have mostly focused on investigating how the bacterial microbiome is
used at a single point in time to predict disease. In this study, we
assessed whether a model built with bacterial microbiome data could
accurately predict adenoma or carcinoma and adjust positive probability
for lesion (adenoma or carcinoma) after surgical treatment. This model
was tested on a \texttt{66} person group that included samples before
and after treatment to allow for the assessment of how the model adjusts
risk after treatment. The model chosen had a 10-fold cross validated AUC
of \texttt{0.819} and a test set AUC of \texttt{0.815}. : For the follow
up samples our Random Forest model significantly decreased the positive
probability of lesion compared to the initial samples for both adenoma
(P-value = \texttt{8.48e-07}) and carcinoma (P-value =
\texttt{5.96e-08}). The top 5 most important operational taxonomic units
(OTUs) for prediction of lesion classified to
\texttt{Lachnospiraceae\ (Otu000013)},
\texttt{Escherichia/Shigella\ (Otu000018)},
\texttt{Ruminococcaceae\ (Otu000020)},
\texttt{Ruminococcus\ (Otu000017)}, and
\texttt{Porphyromonas\ (Otu000153)}. In the test set, the prediction of
lesion for the initial samples had a sensitivity of \texttt{98.5}\%
(\texttt{65/66}) while follow up it was \texttt{100}\%
(\texttt{1\ /\ 1}). Overall, this model could accurately predict those
with an adenoma or carcinoma and lower probability of having a lesion
based on follow up samples after treatment.
